---
title: "Homework 6"
author: "Rebecca Silva"
date: "11/15/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)


knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


### Problem 1 

```{r, warning= FALSE, message=FALSE}
# data read in 
birthweight = 
  read_csv("./data/birthweight.csv") %>% 
  mutate(babysex = factor(babysex, 
                          labels = c("male", "female")), 
         frace = factor(frace, 
                        labels = c("white", "black", "asian", "puerto rican", "other")),
         malform = factor(malform, 
                          labels = c("absent", "present")), 
         mrace = factor(mrace, 
                        labels = c("white", "black", "asian", "puerto rican")))

# 0 missing values
missing_vals = sum(is.na(birthweight))
```

Since we are starting off with 20 variables, we first eliminate some based on general hypotheses about variables that might not explain birthweight well and based on their respective correlation to birthweight `wgt`, shown below.   

```{r, warning = FALSE}
# correalation matrix 
birthweight_numeric = 
  select_if(birthweight, is.numeric)
round(cor(birthweight_numeric %>% select(-bwt), birthweight %>% pull(bwt)), 2)
```


In order to predict birthweight, we want to understand how our given variables relate to birthweight, and ultimately, include the most important predictors of birthweight in our model. We will focus on the variables, `bhead`, `blength`, `gaweeks`, `wtgain`, `delwt`, `mrace`, `frace`, `babysex`, and `smoken`. 

We include some graphical depictions of the relationships that seem to have the strongest association. A linear relationship looks likely between many of the variables and `birthweight`. 
```{r, warning = FALSE}
# baby’s head circumference at birth (centimeters)
birthweight %>% 
  ggplot(aes(x = bhead, y = bwt)) + 
  geom_point(alpha = .5)

# baby's gestational age
birthweight %>% 
  ggplot(aes(x = gaweeks, y = bwt)) + 
  geom_point(alpha = .5)
```


```{r}
mod_full = lm(bwt ~ ., data = birthweight)

# adding all continuous w strong correlation
# and add all catergorical except malform
mod1 = 
  lm(bwt ~ bhead + blength + gaweeks + wtgain + delwt + mrace + frace + babysex  + smoken,
     data = birthweight)
# mod1 %>% broom::glance()
# mod1 %>% broom::tidy()


# add interactions for sex 
mod_sex = 
  lm(bwt~ (bhead + blength + gaweeks + wtgain + delwt + mrace + frace +  + smoken)*babysex,
     data = birthweight)

# keep significant interactions
mod_sex_sig = 
  lm(bwt~ bhead*babysex + blength + gaweeks + wtgain + delwt + smoken+ mrace + frace, 
     data = birthweight)

```

Although we fit a full model at first, we do not use p-values to take out predictors and only use the full fit to compare R-squared measures. After starting off with the predictors mentioned above, we add an interaction term of sex with all the variables and see which ones are significant. This is based on the hypothesis that the levels of sex may effect certain predictor's relationship to birthweight. Since we find that only `bhead*babysex` is significant we keep only that interaction term. 

We next checked for multicoliniearity, and take out father's race since we already include mother race and the two predictors are highly correlated. 

```{r}
# check for multicollinearity 
car::vif(mod_sex_sig)

# take out father's race bc high correlation with mom race 
mod_reduced = lm(bwt~ bhead+ babysex + blength + gaweeks + wtgain + delwt + mrace + smoken, data = birthweight)
#car::vif(mod_reduced)

```


```{r}
# partial f-tests
mod_small = 
  lm(bwt~ bhead*babysex + blength + gaweeks + wtgain + delwt + mrace  + smoken, 
     data = birthweight)

mod_full = lm(bwt ~ ., data = birthweight)

anova(mod_small, mod_full) 

mod_final = 
  lm(bwt~ bhead*babysex + blength + gaweeks + wtgain + delwt + mrace  + smoken, data = birthweight)

mod_final %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

```

Even though a Nested F- test suggests the full model is more useful (significant F-stat), since the adjusted $R^2$ only increases from my model to a full model by 0.001, and my model is simpler and more interpretable, I will go with the smaller model using `bhead`, `babysex`, `bhead*babysex`, `blength`, `gaweeks`, `wtgain`, `delwt`, `mrace`, and `smoken` as predictors.

Although we still use many predictors (8), we focused on prediction of birthweight here and less on interpretation, so the number of predictors is less of a penalty. 

```{r, warning = FALSE, message= FALSE}
birthweight %>% 
  modelr::add_residuals(mod_final) %>% 
  modelr::add_predictions(mod_final) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() +
  geom_hline(yintercept = 0, color = "red")
```

The residuals vs. fitted values plot shows that residuals are centered around 0 but they vary based on the predicted value, indicating the assumption of constant variance among residuals is not quite met. Since the two other models suggested in this problem also have similar patterns in the residual vs fitted plot, we continue with caution.  

##### Comparison with two other models: 

* birthweight ~ blength + gaweeks

* birthweight ~ bhead * blength * babysex

### Checking residuals 
```{r}
mod1 = lm(bwt ~ blength + gaweeks, data = birthweight)
mod2 = lm(bwt ~ bhead*blength*babysex, data = birthweight)

cv_df = 
  crossv_mc(birthweight, 100) 
cv_df =
  cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

birthweight_cv = 
  cv_df %>% 
  mutate(mod  = map(train, ~lm(bwt~ bhead*babysex + blength + gaweeks + wtgain + delwt + mrace  + smoken, data = .x)),
         mod1  = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
         mod2  = map(train, ~lm(bwt ~ bhead * blength * babysex, data = .x))) %>% 
  mutate(rmse_mod = map2_dbl(mod , test, ~rmse(model = .x, data = .y)),
         rmse_mod1 = map2_dbl(mod1, test, ~rmse(model = .x, data = .y)), 
         rmse_mod2 = map2_dbl(mod2, test, ~rmse(model = .x, data = .y)))
```


```{r}
birthweight_cv %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

The chosen model has smallest rmse which is a good way to check if model has predictions with large variances. It seems that on average the rmse is lowest in my chosen model. 

## Problem 2 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}
# function gets r-squared and log(b1b0)
get_terms = function(model){
  output1 = 
    model %>% 
    broom::glance() %>% 
    select(r.squared)
  
  output2 = 
    model %>% 
    broom::tidy() %>% 
    select(estimate)
  
  return(
    df = tibble(
      r_squared = as.numeric(output1), 
      log_b1b2 = log(output2$estimate[1]*output2$estimate[2])
    ))
}
set.seed(1)
weather_boot = 
  weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x)),
    results = map(models, get_terms)) %>% 
  select(results) %>% 
  unnest(results) 
```

```{r}
# plot distributions
weather_boot %>% 
  ggplot(aes( x = r_squared)) +
  geom_histogram(alpha = .5) +
  xlim(c(.87, .95)) +
  labs(
    title = "Distribution of R-Squared", 
    x = "r-squared"
  ) +
  theme(plot.title = element_text(hjust = 0.5))

weather_boot %>% 
  ggplot(aes( x = log_b1b2)) +
  geom_histogram(alpha = .5) +
  xlim(c(1.9, 2.1)) +
  labs(
    title = "Distribution of log(Beta1*Beta0)", 
    x = "log(b1*b0)"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```

The distribution of $R^2$ is symmetric and unimodel with a sharp peak around 0.912. It seems that there is small variation on its values as most fits between 0.88 and 0.94. 
  
R-Squared 95% CI:
```{r}
round(quantile(pull(weather_boot,r_squared), probs = c(0.025, 0.975)), digits = 3)
```

95% of our bootstrap samples obtained an $R^2$ value between 0.894 and 0.927. Since our bootstrap distribution represents our sampling distribution, we could interpret this as being 95% confident that the r-squared from the linear regression between tmax and tmin is between 0.894 and 0.927. 

The distribution of log(Beta1*Beta0) is symmetric and centered around 2.02.

Log(B1*B0) 95% CI:
```{r}
round(quantile(pull(weather_boot,log_b1b2), probs = c(0.025, 0.975)), digits = 3) 
```

We are 95% confident that the log of the product of B_1 and B_0 lies between 1.97 and 2.06. 

